---
title: "a3_task3_lauren_skube"
author: "Lauren Skube"
date: "10/30/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(janitor)
library(here)
library(kableExtra)
library(lubridate)
```

## Summary 
In this document, I use red and purple urchin data from the Santa Barbara LTER Project to practice analyzing and visualizing data and summary statistics. I also practice version control and project management using GitHub.

```{r}
# read in data using the here() function 
urchins <- read.csv(here("data" , "urchins.csv")) %>% 
  clean_names() # convert column names to lower_snake_case
```

## Section 1 - Part 1: Exploratory data visualization 
Create a histogram and a boxplot to compare distribution sizes of purple and red urchins. 

```{r, message=FALSE, warning=FALSE}
ggplot(data = urchins, aes(x= size)) + 
  geom_histogram(bins = 15) + # Create a histogram and change the number of bins 
  facet_wrap(~ common_name) + # To create a multi-panel plot separated by common name 
  labs(x= "Urchin diameter (cm)") # Update axis labels 

ggplot(data = urchins, aes(x = common_name, y = size)) + 
  geom_boxplot() + # Create a box plot from the urchins data
  labs(x = "Common name" , y = "Urchin diameter (cm)") # Update axis labels 
```

## Section 1 - Part 2: Descriptive statistics table 
Here I create a summary table that contains the mean, median, standard deviation, and sample size for purple and red urchins. 

```{r, message=FALSE, warning=FALSE}
urchins_stats <- urchins %>% 
  group_by(common_name) %>% # Use the group_by() and summarize() combo to create a table of summary statistics for urchin size separated by common name 
  summarize(mean_size = mean(size), 
            median_size = median(size), 
            sd_size = sd(size), 
            n = n())
  kbl(urchins_stats) %>% 
  kable_styling()
  
  # Actual difference in mean urchin diameter 
  6.776489 - 4.469797
  
  # Percent differences in mean urchin diameter 
  (6.776489 - 4.469797)/ ((6.776489+4.469797)/2) * 100
```


## Section 1 - Part 3: Analysis of exploratory graphs and summary statistics 
The graph for purple urchins is left skewed and has a central tendency of 4.47 cm. The graph for red urchins is closer to a normal distribution (symmetric) and has a central tendency of 6.78 cm. The medians of each sample is relatively close to the mean. The actual difference in mean urchin diameter for purple urchins and red urchins is 2.307 cm, and the percent difference is 41.02%.The range of red urchin diameter measurements is larger than the range of purple urchin diameter measurements because the graph for red urchins is wider and more spread along the x-axis (has a greater variance within the group), whereas the graph for purple urchins is more narrow (has a smaller variance within the group). 

## Section 2: Explore purple urchin counts by site and year 

```{r, message=FALSE, warning=FALSE}
urchins_clean <- urchins %>% 
  mutate(date = mdy(date)) # Change the dates to proper format and the class to "Date"
class(urchins_clean$date) # Check the class of the date column

purple_urchins <- urchins_clean %>%  # Create a new subset called "purple_urchins"
  filter(common_name == "Purple Urchin") %>% # Only keep data where the common name is Purple Urchin
  mutate(year = year(date)) %>% # Create a new column called "year" that contains only the year
  mutate(site_full = case_when( # Create a new column called "site_full" that contains the full name of the site 
    site %in% "NAPL" ~ "Naples Reef", 
    site %in% "AQUE" ~ "Arroyo Quemado Reef" , 
    site %in% "CARP" ~ "Carpinteria Reef" , 
    site %in% "MOHK" ~ "Mohawk Reef" , 
    site %in% "IVEE" ~ "Isla Vista Reef" # list the current name first and then what it should be called in the new column following the "~"
  ))
  
```
 
## Section 2: Summary table 
Create a summary table containing the total number of purple urchins observed by site and year starting from the purple_urchins subset 
 
```{r}
purple_summary <- purple_urchins %>% # create a table containing the number of purple urchins per site and year
  count(year, site_full)

ggplot(data = purple_summary, aes(x= year, y = n)) +
  geom_col(aes(fill= site_full), show.legend = FALSE) + 
  facet_wrap(~site_full) +
  labs(x= "Year", 
       y = "Total Count of Purple Urchins")
  
```
 
 ## End of task
 
 
 
 
 
 
 
 
 
 
 
 
 
 
 
 


