---
title: "a3_task4_lauren_skube"
author: "Lauren Skube"
date: "11/1/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(janitor)
library(here)
library(lubridate)
```

## Summary 

```{r}
# Read in the urchins.csv data 
urchins <- read.csv(here("data" , "urchins.csv"))
```

## Wrangle the data 

```{r, message=FALSE, warning=FALSE}
# Create a new subset called napl_purple_urchins
napl_purple_urchins <- urchins %>% 
  clean_names() %>% # Convert all column names to lower_snake_case
  mutate(date = mdy(date)) %>% # Convert date column to "Date" class
  mutate(year = year(date) , # Create a new column containing the year
         month = month(date)) %>%  # Create a new column containing the month

# Filter the urchins data 
  filter(common_name == "Purple Urchin" , # Only include Purple Urchins
         year == 2010, # Only include observation from 2010..
         month == 1, # And from January...
         site == "NAPL") %>% # And from the Naples Reef site 
  group_by(treatment)
  
```

## Exploratory graphs 

```{r, message=FALSE, warning=FALSE}
# Create a quantile-quantile graph to compare the sample to normality
ggplot(data = napl_purple_urchins, aes(sample = size )) + 
  geom_qq() + 
  facet_wrap(~treatment) + # To separate by treatment type
  labs(x = "Urchin diameter (cm)" )
```

```{r}
# Create a histogram to compare the sample to normality 
ggplot(data = napl_purple_urchins, aes(x = size)) + 
  geom_histogram(bins = 12) + 
  facet_wrap(~treatment) 
```

## Takeaways from the graphs:

## Find summary statistics 

```{r, message=FALSE, warning=FALSE}
treatment_stats <- napl_purple_urchins %>% 
  group_by(treatment) %>% 
  summarise(mean_treatment = mean(size), 
            sd_treatment = sd(size), 
            n = n(), 
            min_treatment = min(size), 
            max_treatment = max(size))
```

## Perform a two-sample t-test 
- This will compare the means between the annual and the control groups
- Start by creating vectors for size observations from each sample

```{r, message=FALSE, warning=FALSE}
# Create vectors containing observations for each sample 
control_sample <- napl_purple_urchins %>% 
  filter(treatment == "CONTROL") %>% 
  pull(size) # Create a vector for control group observations

annual_sample <- napl_purple_urchins %>% 
  filter(treatment == "ANNUAL") %>% 
  pull(size) # Create a vector for the "annual" group observations

# Run a two-sample t-test using the vectors as inputs
t.test(annual_sample, control_sample)

  
```

```{r}
# Actual difference between means: 
3.916667-3.725806

# Percent difference between means: 
0.190861/((3.91667 + 3.72586)/2) * 100
```


The mean of the Naples Reef annual sample (where giant kelp is removed once each winter) is 3.92 cm. with a standard deviation of 0.966. The mean of the control sample (where giant kelp is left intact) is 3.73 cm with a standard deviation of 0.867. The actual difference between the two means is 0.191 cm and the percent difference is 4.99%. Without considering the p-value, there doesn't seem to be a very strong difference in means between the two groups. When you include consideration of the p-value, which is 0.2536, it further confirms that it's not very likely these are two groups with different means. 

??? What does the confidence interval tell us here?? ask in office hours! 

## Meaning of the p-value





