---
title: "a3_task3_lauren_skube"
author: "Lauren Skube"
date: "10/30/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(janitor)
library(here)
library(kableExtra)
```

## Summary 


```{r}
# read in data using the here() function 
urchins <- read.csv(here("data" , "urchins.csv")) %>% 
  clean_names() # convert column names to lower_snake_case
```

## Section 1 - Part 1: Exploratory data visualization 
Create a histogram and a boxplot to compare distribution sizes of purple and red urchins. 

```{r, message=FALSE, warning=FALSE}
ggplot(data = urchins, aes(x= size)) + 
  geom_histogram(bins = 15) + # Create a histogram and change the number of bins 
  facet_wrap(~ common_name) + # To create a multi-panel plot separated by common name 
  labs(x= "Urchin diameter (cm)") # Update axis labels 

ggplot(data = urchins, aes(x = common_name, y = size)) + 
  geom_boxplot() + # Create a box plot from the urchins data
  labs(x = "Common name" , y = "Urchin diameter (cm)") # Update axis labels 
```

## Section 1 - Part 2: Descriptive statistics table 
Here I create a summary table that contains the mean, median, standard deviation, and sample size for purple and red urchins. 

```{r, message=FALSE, warning=FALSE}
urchins_stats <- urchins %>% 
  group_by(common_name) %>% # Use the group_by() and summarize() combo to create a table of summary statistics for urchin size separated by common name 
  summarize(mean_size = mean(size), 
            median_size = median(size), 
            sd_size = sd(size), 
            n = n())
```

## Section 1 - Part 3: Analysis of exploratory graphs and summary statistics 





